Score ReLiability? Exploring the role of score reliability for effect size heterogeneity

### 2024-05-28 Manuscript_ScoreReLiability.qmd
text update add gitkeep

---
title: "Manuscript"
format: html
editor: visual
---

@@

@@

m <- matrix(c(1,	.2,	.7,	.1,	.198,
              1,	.2,	.8,	.1,	.228),
            nrow = 2)

@@

names(df) <- c("$E[\\delta]$",
               "$Var[\\delta]$",
               "$E[a]$",
               "$Var[a]$",
               "$Var[\\delta\\_0]$")

@@

However, both parameters, ES and score reliability in their notation, can not be independent parameters by definition. ES is the “true” standardized mean difference, unaffected by sampling or measurement error. The ES is standardized using the total standard deviation, which is inflated by measurement error. Equation 1, demonstrating how Cohen’s d is computed demonstrates how the mean difference is standardized in this case. The ES on the other hand is standardized using the true standard deviation in the sense of CTT.

*\$\$* *= \\frac{MD}{\\sigma_T}\$\$*

In equation 4, is the true standard deviation, the square root of the true score variance found in equation 2. The fact that the true standard deviation parameter is found both in equation 4 and equation 2 demonstrates that ES and score reliability can not be independent variables. Therefore, the basic premise of equation 3 is not fulfilled, as the assumption of variable independence could not be fulfilled.

~~Hunter & Schmidt’s discussion of how score reliability affects heterogeneity in ESs is fatally flawed \[issues with the assumption underlying the equation\].~~

Alternatively, assuming that both the mean differences and pooled standard deviations are no constant values across replications, meaning both carry non-zero heterogeneity, we can describe observed ESs as a random variable following a ratio distribution. Such a ratio distribution is the probability distribution, constructed by dividing one random variable by a second variable: \$Z = \\frac{X}{Y}\$ (reference). If the distribution of the components is known, first order taylor approximation may be used to generate estimates of mean and variance of the ratio variable (reference). Using some additional assumptions, we use the Taylor-estimator of the ratio’s variance to demonstrate how differences in score reliability may affect heterogeneity in ES.

\$\$var\[Z\] \\approx \\frac{var\[X\]}{E\[Y\]\^2} - \\frac{2E\[X\]}{E\[Y\]\^3} cov\[X,Y\] + \\frac{E\[X\]\^2}{E\[Y\]\^4} var\[Y\]\$\$

@@

\$\$var\[Z\] \\approx \\frac{var\[X\]}{E\[Y\]\^2} + \\frac{E\[X\]\^2}{E\[Y\]\^4} var\[Y\]\$\$

@@

\$\$var\[ES\] \\approx \\frac{var\[MD\]}{E\[SD\]\^2} + \\frac{E\[MD\]\^2}{E\[SD\]\^4} var\[SD\]\$\$

Correcting ES for score reliability is technically done by correcting the pooled standard deviation: \$SD_c = \\sqrt{a V}\$ . Since is always between zero and one, in the case of imperfect score reliability, SD_c will always be smaller than total SD. Additionally, if the variance in SD was introduced by differences in score reliability, these would be removed by the attenuation correction. However, there are two parameters that may introduce variance in SD across replications.

Firstly, as discussed, differences in score reliability may introduce variation to SD. Secondly however, the underlying latent variable we attempted to measure may have been not identically distributed across replications. If the populations where the replications are collected from are differently spread out in terms of the variable of interest, true differences in SD would remain. Re-analyses of the ManyLabs and Registered Replication Reports have shown that such differences in true variance across replications are the norm, not the exception. In that case, some variance in SD_c would remain, albeit smaller than the variance in SD. We reformulate equation X in terms of SD_c

\$\$var\[ES_c\] \\approx \\frac{var\[MD\]}{E\[SD_c\]\^2} + \\frac{E\[MD\]\^2}{E\[SD_c\]\^4} var\[SD_c\]\$\$

The implication of this equation is, again, not straightforward. The expected value of SD_c is guaranteed to be smaller than the expected value of uncorrected SD. Similarly, the variance in SD_c is guaranteed to be smaller than the uncorrected variance in SD. The expected value is always placed in the denominator of the function, while the variance is placed in the numerator of the second term in equation X+1. This implies that the variance in uncorrected ES can only be larger than the variance in ES_c, if the reduction in mean SD_c due to the attenuation correction increases the total equation result faster than the reduction in variance of SD_c decreases the total equation result. To illustrate this point, we consider some parameter combinations in table 2.

@@

<table>
<tbody>
<tr class="odd">
<td><p>  </p></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
</tr>
<tr class="even">
<td><p>   </p></td>
<td><p>   </p>
<p>E[MD]</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>Var[MD}</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>E[SD]</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>Var[SD]</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>E[SD_c]</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>Var[SD_c]</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>Var[ES]</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>Var[ES_c]</p>
<p>   </p></td>
<td><p>  </p></td>
</tr>
<tr class="odd">
<td><p>  </p></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
</tr>
<tr class="even">
<td><p>   </p></td>
<td><p>   </p>
<p>1</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.2^2</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>1</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.2^2</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.8</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.1^2</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.08</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.087</p>
<p>   </p></td>
<td><p>  </p></td>
</tr>
<tr class="odd">
<td><p>  </p></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
</tr>
<tr class="even">
<td><p>   </p></td>
<td><p>   </p>
<p>1</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.2^2</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>1</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.2^2</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.8</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.05^2</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.08</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.069</p>
<p>   </p></td>
<td><p>  </p></td>
</tr>
<tr class="odd">
<td><p>  </p></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
</tr>
<tr class="even">
<td><p>   </p></td>
<td><p>   </p>
<p>1</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.2^2</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>1</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.2^2</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.8</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>0</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.08</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.063</p>
<p>   </p></td>
<td><p>  </p></td>
</tr>
<tr class="odd">
<td><p>  </p></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
</tr>
<tr class="even">
<td><p>   </p></td>
<td><p>   </p>
<p>1</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.2^2</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>1</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.2^2</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.9</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.1^2</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.08</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.064</p>
<p>   </p></td>
<td><p>  </p></td>
</tr>
<tr class="odd">
<td><p>  </p></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
</tr>
<tr class="even">
<td><p>   </p></td>
<td><p>   </p>
<p>1</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.2^2</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>1</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.2^2</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.7</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.1^2</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.08</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.124</p>
<p>   </p></td>
<td><p>  </p></td>
</tr>
<tr class="odd">
<td><p>  </p></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
</tr>
<tr class="even">
<td><p>   </p></td>
<td><p>   </p>
<p>1</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.2^2</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>1</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.2^2</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.5</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.1^2</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.08</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.32</p>
<p>   </p></td>
<td><p>  </p></td>
</tr>
<tr class="odd">
<td><p> </p></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
</tr>
</tbody>
</table>

 Table 2 demonstrates that in scenarios, where either score reliability was very high to begin with (.9 in this case), leading to very little change in E\[SD\] to E\[SD_c\] (row 5), or where the remaining variance in SD_c was comparably low (rows 3 and 4), the attenuation correction led to a reduction in ES heterogeneity. In most other scenarios, where score reliability was lower (.7 and .5 in rows 6 and 7) or remaining variance in SD_c sufficiently large (row 2), the attenuation correction actually increased the heterogeneity in ES. Note that in the table, we did not manipulate expected value or variance of MD or SD.

\[implications of this distribution concerning Var\[ES\] and how that relates to Var\[rel\]\]

@@

**Part C**

@@

 Wang et al. (2021) used several measures which may be promising for this study. Out of four measures they use to assess negative and positive emotions, three should fit conditions ii), ii) and iv); positive and negative emotional responses to ten photographs shown in the experimental trial, positive and negative state emotion after viewing the ten photographs and positive and negative anticipated emotions following the study. Over the following paragraphs, we will briefly describe the study design and measures used that are of interest to us. Details regarding the protocol, stimuli or additional measures used may be found at \[…\].

@@

 In the experimental trials, the participants viewed 10, similarly negative, photographs of the COVID-19 pandemic and rated their emotions on each photograph using Likert-style responses to the positive/negative questions. After the experimental trials, individuals responded to five negative items from the Differential Emotions Scale on fear, anger, sadness, distrust, and stress. Similarly, they responded to five positive items from the same scale, on hope, gratitude, love, inspiration, and serenity. Averages of the five positive and negative items respectively were used as measures of state emotions after the experimental trials. Similarly, negative and positive anticipated emotions were assessed, by prefacing a question of “In the next week, to what extent, if at all, do you think you will feel each of the following?” and following up with the same items used to assess state emotions. All items or the measures used may be found in the appendix XXX, or at \[url…\].

@@

<table>
<tbody>
<tr class="odd">
<td><p>  </p></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
</tr>
<tr class="even">
<td><p>   </p></td>
<td><p>   </p>
<p>Measure</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>Valence</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>ES</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>Tau ES</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>ES_c</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>Tau ES_c</p>
<p>   </p></td>
<td><p>  </p></td>
</tr>
<tr class="odd">
<td><p>  </p></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
</tr>
<tr class="even">
<td><p>   </p></td>
<td><p>   </p>
<p>Photograph   Response</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>Positive</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>1.065   (.065)</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.377 (sq   .036)</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>1.219   (.080)</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.468 (sq   .056)</p>
<p>   </p></td>
<td><p>  </p></td>
</tr>
<tr class="odd">
<td><p>  </p></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
</tr>
<tr class="even">
<td><p>   </p></td>
<td><p>   </p>
<p>Negative</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>-.663   (.040)</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.220 (sq .014)</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>-.747   (.047)</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.260 (sq   .019)</p>
<p>   </p></td>
<td><p>  </p></td>
<td></td>
<td></td>
</tr>
<tr class="odd">
<td><p>  </p></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
</tr>
<tr class="even">
<td><p>   </p></td>
<td><p>   </p>
<p>Emotional   State</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>Positive</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.213   (.018)</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.052 (sq   .003)</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.234   (.020)</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.054 (sq   .003)</p>
<p>   </p></td>
<td><p>  </p></td>
</tr>
<tr class="odd">
<td><p>  </p></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
</tr>
<tr class="even">
<td><p>   </p></td>
<td><p>   </p>
<p>Negative</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>-.203   (.020)</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.071 (sq   .003)</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>-.221   (.022)</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.078   (.004)</p>
<p>   </p></td>
<td><p>  </p></td>
<td></td>
<td></td>
</tr>
<tr class="odd">
<td><p>  </p></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
</tr>
<tr class="even">
<td><p>   </p></td>
<td><p>   </p>
<p>Anticipated   Emotion</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>Positive</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.159   (.015)</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>0 (sq   .001)</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.172   (.016)</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>0 (.002)</p>
<p>   </p></td>
<td><p>  </p></td>
</tr>
<tr class="odd">
<td><p>  </p></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
</tr>
<tr class="even">
<td><p>   </p></td>
<td><p>   </p>
<p>Negative</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>-.146   (.020)</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.073 (sq   .003)</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>-.145   (.022)</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.079   (.004)</p>
<p>   </p></td>
<td><p>  </p></td>
<td></td>
<td></td>
</tr>
<tr class="odd">
<td><p> </p></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
</tr>
</tbody>
</table>

>>>>

---
title: "Score ReLiability? Exploring the role of score reliability for effect size heterogeneity"
format: html
editor: visual
---

```{r, include = FALSE}
# library loading and installing as necessary

# relevant libraries required for this script
packages <- c("ggplot2", "here", "metafor", "gridExtra", "knitr", "magrittr", "dplyr")

# check, whether library already installed or not - install and load as needed:
apply(as.matrix(packages), MARGIN = 1, FUN = function(x) {

  pkg_avail <- nzchar(system.file(package = x))   # check if library is installed on system

  if(pkg_avail){
    require(x, character.only = TRUE)             # load the library, if already installed

  }else{
    install.packages(x)                           # install the library, if missing
    require(x, character.only = TRUE)             # load after installation
  }
})

@@

## Analytical Arguments

@@

m <- matrix(c(1,	.2,	.7,	.1,
              1,	.2,	.8,	.1),
            nrow = 2, byrow = T)

@@

names(df) <- c("$E[\\delta]$",
               "$Var[\\delta]$",
               "$E[a]$",
               "$Var[a]$")

df$'$Var[\\delta\\_0]$' <- (df[,3]^2) * df[,2] + (df[,1]^2) * df[,4]

@@

However, both parameters, ES $\delta$ and score reliability $a$ in original notation can not be independent parameters by definition. ES $\delta$ is the “true” standardized mean difference, unaffected by sampling or measurement error. The ES $\delta_0$ is standardized using the total standard deviation, which is inflated by measurement error. Equation 1, demonstrating how Cohen’s d is computed demonstrates how the mean difference is standardized in this case. The ES $\delta$ on the other hand is standardized using the true standard deviation in the sense of CTT.

$$\delta = \frac{\Delta}{\sigma_T}$$

In equation 4, $\sigma_T$ is the true standard deviation, the square root of the true score variance found in equation 2. The fact that the true standard deviation parameter $\sigma_T$ is found both in equation 4 and equation 2 demonstrates that ES $\delta$ and score reliability $a$ can not be independent variables. Therefore, the basic premise of equation 3 is not fulfilled, as the assumption of variable independence could not be fulfilled.

Alternatively, assuming that both the mean differences and pooled standard deviations are no constant values across replications, meaning both carry non-zero heterogeneity, we can describe observed ESs as a random variable following a ratio distribution. Such a ratio distribution is the probability distribution, constructed by dividing one random variable by a second variable: $Z = \frac{X}{Y}$ (reference). If the distribution of the components is known, first order taylor approximation may be used to generate estimates of mean and variance of the ratio variable (reference). Using some additional assumptions, we use the Taylor-estimator of the ratio’s variance to demonstrate how differences in score reliability may affect heterogeneity in ES.

$$var[Z] \approx \frac{var[X]}{E[Y]^2} - \frac{2E[X]}{E[Y]^3} cov[X,Y] + \frac{E[X]^2}{E[Y]^4} var[Y]$$

@@

$$var[Z] \approx \frac{var[X]}{E[Y]^2} + \frac{E[X]^2}{E[Y]^4} var[Y]$$

@@

$$var[ES] \approx \frac{var[MD]}{E[SD]^2} + \frac{E[MD]^2}{E[SD]^4} var[SD]$$

Correcting ES for score reliability is technically done by correcting the pooled standard deviation: $SD_c = \sqrt{a V}$ . Since $a$ is always between zero and one, in the case of imperfect score reliability, $SD_c$ will always be smaller than total $SD$. Additionally, if the variance in $SD$ was introduced by differences in score reliability, these would be removed by the attenuation correction. However, there are two parameters that may introduce variance in $SD$ across replications.

Firstly, as discussed, differences in score reliability may introduce variation to SD. Secondly however, the underlying latent variable we attempted to measure may have been not identically distributed across replications. If the populations where the replications are collected from are differently spread out in terms of the variable of interest, true differences in $SD$ would remain. Re-analyses of the ManyLabs and Registered Replication Reports have shown that such differences in true variance across replications are the norm, not the exception. In that case, some variance in $SD_c$ would remain, albeit smaller than the variance in $SD$. We reformulate equation X in terms of $SD_c$

$$var[ES_c] \approx \frac{var[MD]}{E[SD_c]^2} + \frac{E[MD]^2}{E[SD_c]^4} var[SD_c]$$

The implication of this equation is, again, not straightforward. The expected value of $SD_c$ is guaranteed to be smaller than the expected value of uncorrected $SD$. Similarly, the variance in $SD_c$ is guaranteed to be smaller than the uncorrected variance in $SD$. The expected value is always placed in the denominator of the function, while the variance is placed in the numerator of the second term in equation X+1. This implies that the variance in uncorrected ES can only be larger than the variance in $ES_c$, if the reduction in mean $SD_c$ due to the attenuation correction increases the total equation result faster than the reduction in variance of $SD_c$ decreases the total equation result. To illustrate this point, we consider some parameter combinations in table 2.

@@

m2 <- matrix(c(1,	".2^2",	1,	".2^2",	.8,	".1^2",
               1,	".2^2",	1,	".2^2",	.8,	".05^2",
               1,	".2^2",	1,	".2^2",	.8,	0,
               1,	".2^2",	1,	".2^2",	.9,	".1^2",
               1,	".2^2",	1,	".2^2",	.7,	".1^2",
               1,	".2^2",	1,	".2^2",	.5,	".1^2"),
            nrow = 6, byrow = T)

m_num <- matrix(c(1,	.2^2,	1,	.2^2,	.8,	.1^2,
                  1,	.2^2,	1,	.2^2,	.8,	.05^2,
                  1,	.2^2,	1,	.2^2,	.8,	0,
                  1,	.2^2,	1,	.2^2,	.9,	.1^2,
                  1,	.2^2,	1,	.2^2,	.7,	.1^2,
                  1,	.2^2,	1,	.2^2,	.5,	.1^2),
                nrow = 6, byrow = T)

df2 <- as.data.frame(m2)

names(df2) <- c("$E[\\Delta]$",
                "$Var[\\Delta]$",
                "$E[SD]$",
                "$Var[SD]$",
                "$E[SD_c]$",
                "$Var[SD_c]$")

df2$'$Var[\\delta_0]$' <- ((m_num[,2]) / (m_num[,3]^2)) + (((m_num[,1]^2) / (m_num[,3]^4) )* m_num[,4])
df2$'$Var[\\delta]$' <- ((m_num[,2]) / (m_num[,5]^2)) + (((m_num[,1]^2) / (m_num[,5]^4) )* m_num[,6])

knitr::kable(df2, digits = 3)

Table 2 demonstrates that in scenarios, where either score reliability was very high to begin with ($E[a] = .9$ in this case), leading to very little change in $E[SD]$ to $E[SD_c]$ (row 5), or where the remaining variance in $SD_c$ was comparably low (rows 3 and 4), the attenuation correction led to a reduction in ES heterogeneity. In most other scenarios, where score reliability was lower (.7 and .5 in rows 6 and 7) or remaining variance in $SD_c$ sufficiently large (row 2), the attenuation correction actually increased the heterogeneity in ES. Note that in the table, we did not manipulate expected value or variance of MD or SD.

@@

## Empirical Example

@@

Wang et al. (2021) used several measures which may be promising for this study. Out of four measures they use to assess negative and positive emotions, three should fit conditions ii), ii) and iv); positive and negative emotional responses to ten photographs shown in the experimental trial, positive and negative state emotion after viewing the ten photographs and positive and negative anticipated emotions following the study. Over the following paragraphs, we will briefly describe the study design and measures used that are of interest to us. Details regarding the protocol, stimuli or additional measures used may be found at […].

@@

In the experimental trials, the participants viewed 10, similarly negative, photographs of the COVID-19 pandemic and rated their emotions on each photograph using Likert-style responses to the positive/negative questions. After the experimental trials, individuals responded to five negative items from the Differential Emotions Scale on fear, anger, sadness, distrust, and stress. Similarly, they responded to five positive items from the same scale, on hope, gratitude, love, inspiration, and serenity. Averages of the five positive and negative items respectively were used as measures of state emotions after the experimental trials. Similarly, negative and positive anticipated emotions were assessed, by prefacing a question of “In the next week, to what extent, if at all, do you think you will feel each of the following?” and following up with the same items used to assess state emotions. All items or the measures used may be found in the appendix XXX, or at [url…].

@@

```{r, echo = FALSE}
pos_SMDs.df <- read.csv(file = here("Data/Processed/SMDs_positive.csv"))
neg_SMDs.df <- read.csv(file = here("Data/Processed/SMDs_negative.csv"))
pos_state_SMDs.df <- read.csv(file = here("Data/Processed/SMDs_state_positive.csv"))
neg_state_SMDs.df <- read.csv(file = here("Data/Processed/SMDs_state_negative.csv"))
pos_antc_SMDs.df <- read.csv(file = here("Data/Processed/SMDs_antc_positive.csv"))
neg_antc_SMDs.df <- read.csv(file = here("Data/Processed/SMDs_antc_negative.csv"))

### photograph responses

# fit a meta-analytic model to estimates of uncorrected Cohen's d, for positiveness-ratings
pos_SMDs_rma_raw <- metafor::rma(yi = d_raw, 
                                 sei = SE_d_raw, 
                                 data = pos_SMDs.df,
                                 measure = "GEN",
                                 method = "REML")

# fit a meta-analytic model to estimates of corrected Cohen's d, for positiveness-ratings
pos_SMDs_rma_cor <- metafor::rma(yi = d_cor, 
                                 sei = SE_d_cor, 
                                 data = pos_SMDs.df,
                                 measure = "GEN",
                                 method = "REML")

# fit a meta-analytic model to estimates of uncorrected Cohen's d, for negativeness-ratings
neg_SMDs_rma_raw <- metafor::rma(yi = d_raw, 
                                 sei = SE_d_raw, 
                                 data = neg_SMDs.df,
                                 measure = "GEN",
                                 method = "REML")

# fit a meta-analytic model to estimates of corrected Cohen's d, for negativeness-ratings
neg_SMDs_rma_cor <- metafor::rma(yi = d_cor, 
                                 sei = SE_d_cor, 
                                 data = neg_SMDs.df,
                                 measure = "GEN",
                                 method = "REML")

### state responses

# fit a meta-analytic model to estimates of uncorrected Cohen's d, for positiveness-ratings
pos_state_SMDs_rma_raw <- metafor::rma(yi = d_raw, 
                                 sei = SE_d_raw, 
                                 data = pos_state_SMDs.df,
                                 measure = "GEN",
                                 method = "REML")

# fit a meta-analytic model to estimates of corrected Cohen's d, for positiveness-ratings
pos_state_SMDs_rma_cor <- metafor::rma(yi = d_cor, 
                                 sei = SE_d_cor, 
                                 data = pos_state_SMDs.df,
                                 measure = "GEN",
                                 method = "REML")

# fit a meta-analytic model to estimates of uncorrected Cohen's d, for negativeness-ratings
neg_state_SMDs_rma_raw <- metafor::rma(yi = d_raw, 
                                 sei = SE_d_raw, 
                                 data = neg_state_SMDs.df,
                                 measure = "GEN",
                                 method = "REML")

# fit a meta-analytic model to estimates of corrected Cohen's d, for negativeness-ratings
neg_state_SMDs_rma_cor <- metafor::rma(yi = d_cor, 
                                 sei = SE_d_cor, 
                                 data = neg_state_SMDs.df,
                                 measure = "GEN",
                                 method = "REML")

### anticipated respones

# fit a meta-analytic model to estimates of uncorrected Cohen's d, for positiveness-ratings
pos_antc_SMDs_rma_raw <- metafor::rma(yi = d_raw, 
                                       sei = SE_d_raw, 
                                       data = pos_antc_SMDs.df,
                                       measure = "GEN",
                                       method = "REML")

# fit a meta-analytic model to estimates of corrected Cohen's d, for positiveness-ratings
pos_antc_SMDs_rma_cor <- metafor::rma(yi = d_cor, 
                                       sei = SE_d_cor, 
                                       data = pos_antc_SMDs.df,
                                       measure = "GEN",
                                       method = "REML")

# fit a meta-analytic model to estimates of uncorrected Cohen's d, for negativeness-ratings
neg_antc_SMDs_rma_raw <- metafor::rma(yi = d_raw, 
                                       sei = SE_d_raw, 
                                       data = neg_antc_SMDs.df,
                                       measure = "GEN",
                                       method = "REML")

# fit a meta-analytic model to estimates of corrected Cohen's d, for negativeness-ratings
neg_antc_SMDs_rma_cor <- metafor::rma(yi = d_cor, 
                                       sei = SE_d_cor, 
                                       data = neg_antc_SMDs.df,
                                       measure = "GEN",
                                       method = "REML")
